### 2019-12-02 Table1.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Table 1. The research process stages (adapted from Chalmers \& Glasziou, 2009), examples of potential research waste and how ecology and conservation can limit these.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Table 1. The research process stages (adapted from Chalmers \& Glasziou,
2009), examples of potential research waste and how ecology and
conservation can limit these.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\begin{tabular}{>{\bfseries}lll}
\toprule
Research stage & Examples of potential for research waste & Where ecology and conservation can reduce waste\\
\midrule
 & Irrelevant questions asked & Co-development of research questions with stakeholders and using appropriate methodology such as Delphi exercises to avoid issues such as group think or not including the right group of experts or stakeholders\\

\multirow{-2}{*}{\raggedright\arraybackslash Questions relevant to stakeholders} & Previous knowledge not properly taken into account & Make use of evidence synthesis methods (e.g. cumulative meta-analysis, systematic mapping, systematic reviews, meta-analysis) to identify questions that are not satisfactorily answered\\
\cmidrule{1-3}
 & Study poorly designed, under-powered (or over-powered. etc.) & Use simulations or power-analysis prior to undertaking data collection. Predefine effect size of interest with stakeholders (i.e. do not rely on rules of thumb for “statistical significance”)\\

 & Using inappropriate statistical tools (including overfitting etc.) & Better training of early-career researchers in methods. Open code and data to ensure reproducibility of methods\\

 &  & Open science (open methods and data, reproducible methods, sharing code, etc.)\\

\multirow{-4}{*}{\raggedright\arraybackslash Appropriate design and methods} & \multirow{-2}{*}{\raggedright\arraybackslash Questionable research practices3 lead to poor quality research} & Better training of early-career researchers in methods of open science and evidence synthesis\\
\cmidrule{1-3}
 & Lack of open data & Open science (open methods and data, reproducible methods, sharing code, etc.)\\

 & Hypothesising after the results are known & Pre-registration of hypotheses\\

 & p-hacking & Open science (open methods and data, reproducible methods, sharing code, etc.)\\

 & File-drawer syndrome (only some studies are published) & Pre-registration of hypotheses and methods. Open publishing (including preprints)\\

\multirow{-5}{*}{\raggedright\arraybackslash Unbiased reporting} & Incomplete reporting, making evidence synthesis difficult or impossible & Increasing knowledge of researchers and peer reviewers on what is essential to report, and changing journal guidelines where necessary to ensure all relevant information is reported\\
\cmidrule{1-3}
Accessible full publication & Publications not available to practitioners and decision makers & Open access publishing, including making resources available to researchers to be able to publish open access\\
\cmidrule{1-3}
Evidence synthesis & Research not designed or presented in the context of the existing knowledge & Using systematic reviews, systematic maps, meta-analysis, etc. to shape research priorities. Where good quality evidence is available these should be synthesised providing evidence to relevant stakeholders. Research gaps should be the focus of primary studies.\\
\bottomrule
\end{tabular}

\end{document}