### 2020-01-27 Manuscript/SuppMat.Rmd

---
title: "Evidence accumulation for tackling research waste (Supplement)"
csl: nature.csl
output:
  pdf_document:
    fig_caption: yes
  word_document: 
    fig_caption: yes
  html_document: 
    fig_caption: yes
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Matthew J. Grainger^1^*, Friederike C. Bolam^2^, Gavin B. Stewart^2^, Erlend B. Nilsen^1^

^1^Norwegian Institute for Nature Research, P.O. Box 5685 Torgarden, 7485, Trondheim, Norway

^2^Modelling Evidence and Policy Research Group, School of Natural and Environmental Sciences, Ridley Building 2, Claremont Road, Newcastle upon Tyne NE1 7RU, UK

*corresponding author: matthew.grainger\@nina.no

## Supplementary Methods

We extracted the data and R code from the supplementary information in [@darras2018comparing] to recreate their analysis. As such we are dependent on the accurate extraction of data from the primary studies by the original authors. We intended our re-analysis of their data to be an example of the cumulative meta-analysis approach rather than to make explicit recommendations about the use of acoustic recorders for avian surveys. Building on their random effects meta-analysis we ran a cumulative meta-analysis using the “cumul” function in the “metafor” [@metafor] package in R. The cumulative meta-analysis was ordered by publication year and plotted using the ggplot2 package in R[@ggplot2]. Where a single study provided more than one estimate of effect the order in which the estimates were accumulated was the same as the order presented by [@darras2018comparing] and treated as subsequent trials. Changing the order of that these particular studies were accumulated made no difference to the stability of the estimates over time (see the code for an assessment). The original code, the original data, our additional code for running the analysis and plotting can be found at https://github.com/DrMattG/Research_waste.

```{r flowc, echo=FALSE, message=FALSE, warning=FALSE, fig.cap= "The production of research flows through five stages (blue lines) all of which can lead to research waste (red dashed lines). Ecology and conservation have begun to reduce waste by focusing on methodological improvements and open science. Evidence synthesis (including reporting to decision makers) can contribute to the reduction in research waste by influencing question setting and appropriate methods and design (black dashed lines). Poor evidence synthesis can also lead to research waste. \\label{flow}"}

library(DiagrammeR)

graph<-grViz("
    digraph boxes_and_circles {
    # a 'graph' statement
    graph [overlap = true, fontsize = 10]

    # left 'node' statements
    node [shape = box, fontname = Helvetica]
      tab1 [label = '@@1',color = blue]
      tab2 [label = '@@2',color = blue]
      tab3 [label = '@@3',color = blue]
      tab4 [label = '@@4',color = blue]
      tab6 [label = '@@6',color = blue]

    # right node statements
      tab8 [label = '@@8',color = red]
      tab9 [label = '@@9',color = red]
      tab10 [label = 'Lack of transparency',color = red]
      tab11 [label = 'Publications not available to practitioners \\n& decision makers',color = red]

      node [shape = circle,fixedsize = true,
        width = 2, color=blue] // sets as circles
    tab7 [label = '@@7']

    node [shape = circle,
        fixedsize = true,
        width = 2, color=red] // sets as circles
    tab5 [label = '@@5']

        # several 'edge' statements
 tab1->tab2->tab3->tab4->tab6->tab7 [color=blue]
 tab1->tab8 [color=red,style=dashed]
  tab2->tab9 [color=red,style=dashed]
   tab3->tab10 [color=red,style=dashed]
    tab4->tab11 [color=red,style=dashed]
    tab8->tab5 [color=red,style=dashed]
    tab9->tab5 [color=red,style=dashed]
    tab10->tab5 [color=red,style=dashed]
    tab11->tab5 [color=red,style=dashed]
    tab6->tab1 [style=dashed]
        tab6->tab2 [style=dashed]
    tab6->tab5 [color=red, style=dashed]

    tab8->tab9[style=invis]
    tab9->tab10[style=invis]
    tab10->tab11[style=invis]

        # define ranks

         subgraph {
            rank = same; tab6; tab11
         }

    }
       [1]: 'Questions are relevant to stakeholders'
      [2]: 'Appropriate design & methods used'
      [3]: 'Unbiased reporting'
      [4]: 'Accessible full publication'
      [5]: 'Wasteful research'
      [6]: 'Evidence synthesis'
      [7]: 'Utilisable research' 
      [8]: 'Previous knowledge not accounted for'
      [9]: 'Poor design; \\nquestionable research practices'
        ")#, height=1000)

graph

```

# References {-}

<div id="refs"></div>

\newpage